---
title: "PML Assignment: Prediction Writeup"
keep_md: yes
---

## Load library and set seed
We load the library and set seed.
```{r, message=FALSE}
library(caret)
library(doParallel)
set.seed(123)
```

## Read and clean data
We read data from the csv files. Note there is header in both file.
```{r}
train_raw <- read.csv("pml-training.csv", header = TRUE, na.strings = c("NA", ""))
test_raw <- read.csv("pml-testing.csv", header = TRUE, na.strings = c("NA", ""))
```

Next clean the data. Remove NA value. Remove unwanted columns.
```{r}
train_clean <- train_raw[, (colSums(is.na(train_raw)) == 0)]
test_clean <- test_raw[, (colSums(is.na(test_raw)) == 0)]
train_clean_r <- train_clean[,-(1:7)]
test_clean_r <- test_clean[,-(1:7)]
```

## Model Building
We split the dataset into 70% training dataset and 30% validation.
```{r}
inTrain = createDataPartition(train_clean_r$classe, p=.7, list = FALSE)
training = train_clean_r[inTrain,]
testing = train_clean_r[-inTrain,]
```

## PCA
We might not need every predictor. Apply PCA (exclude classe in this).
```{r}
preProc <- preProcess(training[, -53], method = "pca")
trainPreP <- predict(preProc, training[, -53])
testPreP <- predict(preProc, testing[, -53])
```

## Train model
We train the model using random forest with ntree and mtry fixed after trial & error (considering runtime versus accuracy).
```{r, message=FALSE}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
modelFit <- train(training$classe ~ ., method = "rf", data = trainPreP,ntree=100, tuneGrid=data.frame(mtry=2))
stopCluster(cl)
```

## Cross Validation
We apply the train model to cross validation set. We view result using confusionMatrix. 
```{r}
pred_VRF <- predict(modelFit, testPreP)
confusionMatrix(testing$classe, pred_VRF)
```
Expected Out of sample error is 1-Accuracy, which is **0.0274**.

## Predicted results
We preprocessed the original testing set and also removed the "problem_id" column.
Running our model with test data and we get the predicted results
```{r}
testFinal <- predict(preProc, test_clean_r[, -53])
predfinal <- predict(modelFit, testFinal)
predfinal
```




